Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-21 paper/sections/proof-reconstruction.tex
'.

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
In the \Athena analysis, some unnecessary steps that introduce redundancies
and some unused input is removed from the proof-tree.

As a result of using \Athena, we get an \Agda file of the proof.
To generate this file, \Athena uses names of functions and theorems from
our \Agda libraries: \texttt{agda-prop}, and \texttt{agda-metis}.

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the proposed formalizations for the
propositional logic and the \Metis inferences.
In that case \Agda outputs an interface file.

Otherwise, when type-checking fails, the failure must be
investigated by the user looking at the error in the \TSTP derivation by \Metis,
in the translation by \Athena,in the \Agda formalisations mentioned above or
even in the type-checker, \Agda..

@@

The type of natural numbers is called \Nat type, and it is defined as
usual, i.e., \zero and \suc are the data constructors. We use names
and symbols of the arithmetic operations as usual.
We include syntax sugar for $\zero$, $\suc$, $\suc(\zero)$,
$\cdots$, with the decimal representation: $0$, $1$, $2$, $\cdots$
as well.

@@

In this section, we emulate each \Metis inference rule with
function and its respective theorem.
The following pattern is used to present the formalization.
The function name written in bold face
is a \Metis inference rule in \TSTP derivations.
The same function name written in roman letters refers to the proposed
version of the rule.
Each rule proposed has a respective theorem also written in roman letters
that proves its soundness.
This theorem is named by using the prefix \emph{thm}
and the name of the rule.

@@

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}
\item \rm{thm-rule} is the theorem that proves the soundness of the
$\rm{rule}$ function.
\begin{equation*}\text{If }Γ ⊢ φ\text{ then }Γ ⊢ rule\ φ.\end{equation*}

@@

\begin{definition}[strip]
\label{def:strip}
  \begin{align}
  \label{eq:strip}
    &\fstrip &&: \Prop → \Nat → \Prop\\
    &\fstrip\ (φ₁ ∧ φ₂)\ n     &&= \fuh\ (\fstrip\ φ₁) ∧ \fuh\ (φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ∨ φ₂)\ n     &&= \fuh\ (¬ φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ⇒ φ₂)\ n     &&= \fuh\ (φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ⇔ φ₂)\ n     &&= \fuh\ φ₁ ⇒ \fstrip\ φ₂) ∧ \fuh\ φ₂ ⇒ \fstrip φ₁)\\
    &\fstrip\ (¬ (φ₁ ∧ φ₂))\ n &&= \fuh\ φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ∨ φ₂))\ n &&= \fuh\ \fstrip(¬ φ₁)) ∧\ \fuh ¬ φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ⇒ φ₂))\ n &&= \fuh\ \fstrip(φ₁)) ∧\ \fuh φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ⇔ φ₂))\ n &&= \fuh\ φ₁ ⇒ \fstrip(¬ φ₂)) ∧\ \fuh ¬ φ₂ ⇒ \fstrip(φ₁))\\
    &\fstrip\ (¬ ¬ φ₁)\ n      &&= \fuh\ \fstrip(φ₁))\\
    &\fstrip\ (¬ ⊥)\ n         &&= ⊤\\
    &\fstrip\ (¬ ⊤.)\ n        &&= ⊥\\
    &\fstrip\ φ\ n             &&= φ
  \end{align}
\end{definition}

@@

\begin{lemma}[lem-inv-strip]
If $Γ ⊢ \fstrip\ φ$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

@@

\begin{itemize}
\item Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip(¬¬φ₁,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(\fstrip(φ₁,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip(φ₁,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁)$}
  \RightLabel{lem-inv-uh.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ φ₂$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip(φ, n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁)$}
\RightLabel{lem-inv-uh,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}

  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

  \AxiomC{$\fstrip(φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip(φ)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

In Def.~\ref{def:resolve}, the emulated rule for $resolve$
has four arguments. The first three were described
above, the last one, ψ, represents the expected conclusion in the \Metis
derivation.

@@

% φ ⊻ ψ if and only if ¬ (φ ⇔ ψ).

@@

Now, we have removed redundancies in the disjunctions by applying
$\frm_{∨}$ function, we can define a similar function to work with
conjunctions, i.e., $\frm_{∧}$.

@@

\begin{lemma}[lem-nnf]
  \label{lem:lem-nnf}
  If $Γ ⊢ φ$ then $Γ ⟝ \fnnf(φ)$ where,
  \begin{align*}
   \begin{split}
     &\fnnf : \NProp \to \NProp\\
     &\fnnf(φ) = \fnnf₀(⊕, φ).
   \end{split}
  \end{align*}
\end{lemma}

@@

\begin{definition}[canonicalize]
  \label{def:canonicalize}
  \begin{align*}
    \begin{split}
      &\fcanonicalize : \Prop \to \Prop \to \Prop\\
      &\fcanonicalize(φ, ψ) =
      \begin{cases}
        ψ, &\text{ if }ψ ≡ φ\\
        ψ, &\text{ if }ψ ≡ \fform(\fnnf(φ))\\
        ψ, &\text{ if }\fcnf(ψ) ≡ \freorder_{∧∨}(\fcnf(φ), \fcnf(ψ))\\
        φ, &\text{ otherwise.}\\
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

@@

\begin{proof}\hspace{3mm}
\begin{itemize}
\item If $φ ≡ ψ$ by substitution theorem we conclude $Γ ⊢ ψ$.
\item If $ψ ≡ \fform(\fnnf(φ))$,
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{lem-nnf}
    \UnaryInfC{$Γ ⊢₂~\fnnf(φ)$}
    \RightLabel{lem-form}
    \UnaryInfC{$Γ ⊢ \fform(\fnnf(φ))$}
    \AxiomC{$ψ ≡ \fform(\fnnf(φ))$}
    \RightLabel{subst.}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}

\item If $γ = \fcnf(ψ) ≡ \freorder_{∧∨}(\fcnf(φ), \fcnf(ψ))$,
  \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ$}
      \RightLabel{thm-cnf}
      \UnaryInfC{$Γ ⊢ \fcnf(φ)$}
      \RightLabel{thm-reorder$_{∧∨}$}
      \UnaryInfC{$Γ ⊢ \freorder_{∧∨}(\fcnf(φ), \fcnf(ψ))$}
      \AxiomC{$γ$}
      \RightLabel{subst}
      \BinaryInfC{$Γ ⊢ \fcnf(ψ)$}
      \RightLabel{thm-inv-cnf.}
      \UnaryInfC{$Γ ⊢ ψ$}
    \end{bprooftree}
  \end{equation*}
\end{itemize}
\end{proof}

@@

\begin{definition}[clausify]
  \label{def:clausify}
 \begin{align*}
   \begin{split}
      &\fclausify : \Prop → \Prop → \Prop\\
      &\fclausify(φ,ψ) =
      \begin{cases}
        ψ, &\text{ if }φ≡ψ\\
        \freorder_{∧∨}(\fcnf(φ), ψ), &\text{ otherwise.}
      \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-clausify]
\label{thm:thm-clausify}
  $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fclausify(φ, ψ)$.
\end{theorem}

>>>>

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
In the \Athena analysis, some unnecessary steps that introduce redundancies
and some unused input is removed from the proof-tree.
As result, we get from \Athena an \Agda file of the proof
with names of functions and theorems from
the \Agda libraries that accompany this article:
\texttt{agda-prop} and \texttt{agda-metis}.

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the proposed formalizations for the
propositional logic and the \Metis inferences.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user looking at the error in the \TSTP derivation
by \Metis, in the translation by \Athena, in the \Agda formalizations
mentioned above or even in the type-checker~\Agda.

@@

The type of natural numbers is called \Nat type, and it is defined as
usual, \ie \zero and \suc are the data constructors. We use names
and symbols of the arithmetic operations as usual.
We include syntax sugar for $\zero$, $\suc$, $\suc(\zero)$,
$\cdots$, with the decimal representation: $0$, $1$, $2$, $\cdots$
as well.

@@

In this section, we emulate each \Metis inference rule with a
function and its respective theorem.
The following pattern is used to present the formalization.
The function name written in bold font
is a \Metis inference rule in \TSTP derivations.
The same function name written in roman letters refers to the proposed
version of the rule.
Each rule proposed has a respective theorem also written in roman letters
that proves its soundness.
This theorem is named by using the prefix \emph{thm}
and the name of the rule.

@@

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}
\item \rm{thm-rule} is the theorem that proves the soundness of the
$\rm{rule}$ function.
\begin{equation*}\text{If }Γ ⊢ φ\text{ then }Γ ⊢ \rm{rule}\ φ.\end{equation*}

@@

\begin{definition}[strip]
\label{def:strip}
\label{eq:strip}

\begin{equation}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     ~(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh (¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh (φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      ~(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         ~(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         ~(\suc~n) &= ⊥\\
\fstrip &φ             ~n &= φ
\end{array}
\end{aligned}
\end{equation}
\end{definition}

@@

\begin{lemma}[lem-inv-strip]
If $Γ ⊢ \fstrip~φ$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

@@

\begin{itemize}
\item Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ φ₂$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(φ)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip~φ₂~n$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~φ, n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip~φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁)$}
\RightLabel{lem-inv-uh,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~φ,n-1)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip~φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁ ⇒ \fstrip~φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

  \AxiomC{$\fstrip~φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip~φ)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

In Def.~\ref{def:resolve}, the proposed rule for $resolve$
has four arguments. The first three were described
above, the last one, ψ, represents the expected conclusion in the \Metis
derivation.

@@

@@

Now, we have removed redundancies in the disjunctions by applying
$\frm_{∨}$ function, we can define a similar function to work with
conjunctions, \ie, $\frm_{∧}$.

@@

\begin{lemma}[lem-nnf]
  \label{lem:lem-nnf}
  If $Γ ⊢ φ$ then $Γ ⟝ \fnnf~φ$ where,
  \begin{align*}
   \begin{split}
     &\fnnf : \NProp \to \NProp\\
     &\fnnf~φ = \fnnf₀(⊕, φ).
   \end{split}
  \end{align*}
\end{lemma}

@@

\begin{definition}[canonicalize]
  \label{def:canonicalize}
  \begin{align*}
    \begin{split}
      &\fcanonicalize : \Prop \to \Prop \to \Prop\\
      &\fcanonicalize(φ, ψ) =
      \begin{cases}
        ψ, &\text{ if }ψ ≡ φ\\
        ψ, &\text{ if }ψ ≡ \fform(\fnnf~φ)\\
        ψ, &\text{ if }\fcnf~ψ ≡ \freorder_{∧∨}(\fcnf~φ, \fcnf~ψ)\\
        φ, &\text{ otherwise.}\\
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

@@

\begin{proof}\hspace{3mm}
\begin{itemize}
\item If $φ ≡ ψ$ by substitution theorem we conclude $Γ ⊢ ψ$.
\item If $ψ ≡ \fform(\fnnf~φ)$,
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{lem-nnf}
    \UnaryInfC{$Γ ⊢₂~\fnnf~φ$}
    \RightLabel{lem-form}
    \UnaryInfC{$Γ ⊢ \fform(\fnnf~φ)$}
    \AxiomC{$ψ ≡ \fform(\fnnf~φ)$}
    \RightLabel{subst.}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}

\item If $γ = \fcnf~ψ ≡ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$,
  \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ$}
      \RightLabel{thm-cnf}
      \UnaryInfC{$Γ ⊢ \fcnf~φ$}
      \RightLabel{thm-reorder$_{∧∨}$}
      \UnaryInfC{$Γ ⊢ \freorder_{∧∨}(\fcnf~φ, \fcnf~ψ)$}
      \AxiomC{$γ$}
      \RightLabel{subst}
      \BinaryInfC{$Γ ⊢ \fcnf~ψ$}
      \RightLabel{thm-inv-cnf.}
      \UnaryInfC{$Γ ⊢ ψ$}
    \end{bprooftree}
  \end{equation*}
\end{itemize}
\end{proof}

@@

\begin{definition}[clausify]
  \label{def:clausify}
 \begin{align*}
   \begin{split}
      &\fclausify : \Prop → \Prop → \Prop\\
      &\fclausify~φ~ψ =
      \begin{cases}
        ψ, &\text{ if }φ≡ψ\\
        \freorder_{∧∨}~(\fcnf~φ~ψ), &\text{ otherwise.}
      \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-clausify]
\label{thm:thm-clausify}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fclausify~φ~ψ$.
\end{theorem}
